---
title: "5 minutes map"
format: gfm
editor: visual
---

This is a tutorial on how to quickly plot data with geographical coordinates on a map. 
## Packages
```{r}
#| message: false
#| warning: false
 
library(rnaturalearth)
library(sf)
library(dplyr)

```

## Base map
Let’s create a ggplot object using the base map to display the data of interest. 
```{r}

uk_map_raw <- ne_countries(
  country = "United Kingdom",
  scale = "large", #small, medium or large
  returnclass = "sf" #sf or sv
)

```


```{r}
library(ggplot2)

uk_map <- ggplot() +
  geom_sf(data = uk_map_raw, 
          fill = "grey95", 
          color = "grey40", 
          linewidth = 0.3) +
  coord_sf(
    xlim = c(-8.8, 2.0),
    ylim = c(49.5, 61.0),
    expand = FALSE
  ) +
  theme_bw()

uk_map
```

Adding a ggplot layer showing regional divisions. 
```{r}
library("maps")

uk_regions <- ne_states(country = "United Kingdom", 
                        returnclass = "sf")

```

```{r}

uk_map_region <- ggplot() +
  
  geom_sf(data=uk_regions, 
          fill = "#eeefefff", 
          color = "white", 
          linewidth = 0.3
          )+ 
  
  geom_sf(data = uk_map_raw, 
          fill = NA, 
          color = "black", 
          linewidth = 0.3) +
  
  coord_sf(
    xlim = c(-8.8, 2.0),
    ylim = c(49.5, 61.0),
    expand = FALSE
  )+
  
  theme_void()

uk_map_region

```

## Plotting data

importing the data to be plotted.
```{r}

data_ids <- read.csv("~/Desktop/5_minutes_map/data_map.csv")

# checking the class of each column
str(data_ids)

```
### Dotplot
Plotting the data, with colours assigned according to the ‘sex’ column.
```{r}

uk_map_region +
  geom_point(data = data_ids, 
             aes(x=longitude, 
                 y=latitude, 
                 color=SEX
                 )
             )
  
  
```

### Heatmap
```{r}

data_ids_heatmap <- as.data.frame(table(data_ids$region))
colnames(data_ids_heatmap) <- c("region", "cases")

data_ids_heatmap <- uk_regions %>%
  left_join(data_ids_heatmap, by = "region")

```

```{r}
#| message: false
#| warning: false

uk_map_region +
  geom_sf(
    data = data_ids_heatmap,
    aes(fill = cases),
    color = NA,
    linewidth = 0.3
  )+
  scale_fill_viridis_c(
    name = "cases",
    option = "magma"
  )

```